Stochastic Variance-Reduced Policy Gradient

### 2018-02-08 svrpg.tex
fixed Lipschitz constant notation

\begin{definition}
For a random variable $X$:
\begin{align*}
        \mathbb{E}_{t\vert s}\left[X\right] &\coloneqq 
		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\vert\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t-1$}}}{\left[X \vert \wt{\vtheta^s}\right]} \\
	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}}]{
			\EVV[\tau_{0,h}\sim\pi(\cdot\vert\vtheta_0^{s+1}) \forall h \in {\it {\bf B}}]
				{\dots
					\EVV[\tau_{t,h}\sim\pi(\cdot\vert\vtheta_t^{s+1})\forall h \in {\it {\bf B}}]
						{X\vert\vtheta_t^{s+1}}
				 \dots
			\vert\vtheta_0^{s+1}}
		\vert\wt{\vtheta}^s}\\
        &= \EVV[t-1|s]{\EVV[\tau_{t,h}\sim \pi(\cdot|\vtheta^{s+1}_t)]{X|\vtheta^{s+1}_t}}
\end{align*}
where the sequence $\wt{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$. 
\todopir{what about $\vtheta^{s+1}_0$ and $\wt{\vtheta}^s$? They are equal. How this affect previous definition?}

@@

\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L_1$.
\end{restatable}
\begin{proof}
By definition of $J(\vtheta)$:
\begin{align}
\Dij{J(\vtheta)}{\theta} 
&= \int_{\Tspace}\Dij{}{\theta}p(\tau|\vtheta)\Reward(\tau)\de \tau
\nonumber\\ 
&= \int_{\Tspace}p(\tau|\vtheta)\score{\vtheta}{\tau}\score{\vtheta}{\tau}^T\Reward(\tau)\de \tau + \int_{\Tspace}\pol(\tau)\Dij{}{\theta}\log p(\tau|\vtheta)\Reward(\tau)\de \tau \nonumber\\
&\leq \sup_{\tau \in \mathcal{T}} \left\{\left|\Reward(\tau)\right|\right\} \left(H^2\GRADLOG^2+H\HESSLOG\right) \label{eq:0}\\
&= \frac{1-\gamma^H}{1-\gamma}RH\left(H\GRADLOG^2+\HESSLOG\right),\nonumber
\end{align}
where \ref{eq:0} is from Assumption \ref{ass:bounded_score}.
Since the Hessian is bounded, $J(\vtheta)$ is Lipschitz-smooth.
\end{proof}

\begin{restatable}[]{lemma}{l-smooth-g}\label{lemma:gsmooth}
Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, $g(\tau\vert\vtheta)$ is Lipschitz continuous with Lipschitz constant $L_2$, \ie for any trajectory $\tau\in\Tspace$:
\[
	\norm[2]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_2\norm[2]{\vtheta-\vtheta'}^2.
\]
\todopir{I would denote $L_J$ and $L_g$ rather than 1 and 2}
\end{restatable}
\begin{proof}
        For both REINFORCE and G(PO)MDP, $g(\tau\vert\vtheta)$ is a linear combination of terms of the kind $\nabla \log \pi_{\vtheta}(a_t\vert s_t)\gamma^t r_t$ \cite{peters2008reinforcement}. These terms have bounded gradient from the second inequality of Assumption \ref{ass:bounded_score} and the fact that $|r_t|\leq R$. Bounded gradient implies Lipschitz continuity. Finally, the linear combination of Lipschitz continuous functions is Lipschitz continuous.
\end{proof}

@@

+\frac{L_2^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}

@@

    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}

@@

	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}

@@

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
	+\frac{f_tW}{B\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets[t-1]{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}, \\
	&\Psi_t = \alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2},
\end{align*}
where $L=\max\left\{L1,L2\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.

>>>>

\begin{definition}
For a random variable $X$:
\begin{align*}
        \mathbb{E}_{t\vert s}\left[X\right] &\coloneqq 
		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}} \\ \tau_{i,h}\sim\pi(\cdot\vert\vtheta^{s+1}_i) \forall h \in {\it {\bf B}}, \text{ for $i=0,\dots,t$}}}{\left[X \vert \wt{\vtheta^s}\right]} \\
	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\vert\wt{\vtheta}^s)\forall j \in {\it {\bf N}}]{
			\EVV[\tau_{0,h}\sim\pi(\cdot\vert\vtheta_0^{s+1}) \forall h \in {\it {\bf B}}]
				{\dots
					\EVV[\tau_{t,h}\sim\pi(\cdot\vert\vtheta_t^{s+1})\forall h \in {\it {\bf B}}]
						{X\vert\vtheta_t^{s+1}}
				 \dots
			\vert\vtheta_0^{s+1}}
		\vert\wt{\vtheta}^s}\\
        &= \EVV[t-1|s]{\EVV[\tau_{t,h}\sim \pi(\cdot|\vtheta^{s+1}_t)]{X|\vtheta^{s+1}_t}}
\end{align*}
where the sequence $\wt{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}, ${\it {\bf N}} = [0,\dots,N)$, and ${\it {\bf B}} = [0,\dots,B)$. To avoid inconsistencies, we also define $\Ets[(-1)]{X} \coloneqq \Es{X}$.

@@

\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
	Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth for some positive Lipschitz constant $L_J$.
\end{restatable}
\begin{proof}
By definition of $J(\vtheta)$:
\begin{align}
\Dij{J(\vtheta)}{\theta} 
&= \int_{\Tspace}\Dij{}{\theta}p(\tau|\vtheta)\Reward(\tau)\de \tau
\nonumber\\ 
&= \int_{\Tspace}p(\tau|\vtheta)\score{\vtheta}{\tau}\score{\vtheta}{\tau}^T\Reward(\tau)\de \tau + \int_{\Tspace}\pol(\tau)\Dij{}{\theta}\log p(\tau|\vtheta)\Reward(\tau)\de \tau \nonumber\\
&\leq \sup_{\tau \in \mathcal{T}} \left\{\left|\Reward(\tau)\right|\right\} \left(H^2\GRADLOG^2+H\HESSLOG\right) \label{eq:0}\\
&= \frac{1-\gamma^H}{1-\gamma}RH\left(H\GRADLOG^2+\HESSLOG\right),\nonumber
\end{align}
where \ref{eq:0} is from Assumption \ref{ass:bounded_score}.
Since the Hessian is bounded, $J(\vtheta)$ is Lipschitz-smooth.
\end{proof}

\begin{restatable}[]{lemma}{l-smooth-g}\label{lemma:gsmooth}
Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, $g(\tau\vert\vtheta)$ is Lipschitz continuous with Lipschitz constant $L_g$, \ie for any trajectory $\tau\in\Tspace$:
\[
	\norm[2]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_g\norm[2]{\vtheta-\vtheta'}^2.
\]
\end{restatable}
\begin{proof}
        For both REINFORCE and G(PO)MDP, $g(\tau\vert\vtheta)$ is a linear combination of terms of the kind $\nabla \log \pi_{\vtheta}(a_t\vert s_t)\gamma^t r_t$ \cite{peters2008reinforcement}. These terms have bounded gradient from the second inequality of Assumption \ref{ass:bounded_score} and the fact that $|r_t|\leq R$. Bounded gradient implies Lipschitz continuity. Finally, the linear combination of Lipschitz continuous functions is Lipschitz continuous.
\end{proof}

@@

+\frac{L_g^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}

@@

    +\frac{L_g^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}

@@

	&\qquad+\frac{L_g^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}

@@

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
	+\frac{f_tW}{B\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets[t-1]{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}, \\
	&\Psi_t = \alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2},
\end{align*}
where $L=\max\left\{L_J,L_g\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.
